Shifting Shadows of the Six: Decoding Toronto's Crime Trends from 2014 to 2023

### 2024-01-25 outputs/paper/paper.qmd
check for grammar

The findings of this paper are noteworthy for several reasons. First, they provide a nuanced understanding of the crime dynamics a large metropolitan area, offering insights that are vital for policymakers, law enforcement, and community leaders. The patterns observed suggest shifts in criminal behavior and possibly reflect broader social and economic changes in the city. Second, the comparative analysis between violent and non-violent crimes offers a unique perspective on the nature of criminal activities, contributing to a more targeted approach in crime prevention and community safety initiatives.

The paper is structured to facilitate a comprehensive understanding of the study and its implications. Following @sec-intro, @sec-data presents the data, detailing the data sources, analytical techniques, and the rationale behind the chosen methods. @sec-result discusses the results, elaborating on the observed trends and patterns in crime data. @sec-discussion provides an in-depth discussion of these findings, exploring potential factors influencing these trends, drawing connections to broader urban issues, and suggestions for future research in this area.

@@

## Data Manipulation and Cleaning
The data was processed and cleaned using R[@citeR], a powerful statistical programming language. For key operations please refer to the [@sec-appendix].

## Variables Explanation
To better understanding the data, summary_table dataset was developed to offer a more structured and aggregated view of the crime data. By transforming and summarizing the data into a format that displays crime types against years, this dataset simplifies the task of identifying and analyzing trends over time. The dataset includes various crime categories, such as assault, theft, and other types, with each variable representing the annual count of reported cases for each crime type. This longitudinal data provides a detailed view of the crime trends in Toronto over a decade. 
```{r}
#| label: tbl-data-summary
#| tbl-cap: Crime Incidents Summary by Year
#| echo: false

# Read the data
summary_table <- read.csv(here("inputs/data/summary_table.csv"), check.names = FALSE)

@@

# Read the data
two_types_crime <- read.csv(here("inputs/data/two_types_crime.csv"), check.names = FALSE)

@@

# Plotting the line graph
ggplot(long_data, aes(x = Year, y = Count, group = Crime_Type, color = Crime_Type)) +
  geom_line(linewidth = 1) + # Use linewidth instead of size
  geom_point(size = 1.5) + # Point size
  theme_minimal() + # Minimal theme
  labs(title = "Crime Count by Type Over Years",
         x = "Year",
         y = "Number of Cases",
         color = "Crime Type") +
  theme(
    plot.title = element_text(hjust = 0.5), # Center the plot title
    legend.title = element_blank(), # Remove the legend title
    axis.text.x = element_text(angle = 45, hjust = 1), # Angle the x-axis text for readability
    text = element_text(size = 16), # Increase the global font size
  )
```
@fig-trends-in-crime reveals distinct trends for various crime types in Toronto over the years. We notice that assault crimes have been on the rise in recent years, this could signal a growing concern in public safety and might prompt law enforcement to investigate the underlying causes, such as increased urbanization, changes in population density, or other socio-economic factors. Conversely, there is a noticeable decline in theft-related crimes, it could reflect the success of recent crime prevention initiatives, improved security measures, or increased public awareness.

@@

* **Violent Crimes (Red Line):** There is an overall increasing trend in the number of violent crime cases over the years. There are some fluctuations, with a noticeable dip occurring in one of the years, but the general direction is upwards, especially towards the end of the period, indicating a rise in violent crimes or an increase in reporting.

@@

## Weaknesses and next steps

While the visualization provides a high-level overview of crime trends in Toronto, it's essential to conduct more detailed analysis to understand the underlying causes of these trends. This includes examining subcategories of crimes, focusing on specific geographic areas within the city, and considering other relevant data sources.

Furthermore, this study emphasises the need of ongoing monitoring and analysis of crime data in comprehending the ever-changing dynamics of urban crime. In light of these findings, future study should focus on understanding the influence of specific policies and social changes on crime rates. Only by doing such a comprehensive analysis can we hope to develop strategies that will ensure the safety and well-being of all Toronto residents.

>>>>

The findings of this paper are noteworthy for several reasons. First, they provide a nuanced understanding of the crime dynamics in a large metropolitan area, offering insights that are vital for policymakers, law enforcement, and community leaders. The patterns observed suggest shifts in criminal behavior and possibly reflect broader social and economic changes in the city. Second, the comparative analysis between violent and non-violent crimes offers a unique perspective on the nature of criminal activities, contributing to a more targeted approach in crime prevention and community safety initiatives.

The paper is structured to facilitate a comprehensive understanding of the study and its implications. Following @sec-intro, @sec-data presents the data, detailing the data sources, analytical techniques, and the rationale behind the chosen methods. @sec-result discusses the results, elaborating on the observed trends and patterns in crime data. @sec-discussion provides an in-depth discussion of these findings, exploring potential factors influencing these trends, drawing connections to broader urban issues, and providing suggestions for future research in this area.

@@

## Data Manipulation and Cleaning
The data was processed and cleaned using R[@citeR], a powerful statistical programming language. For key operations, please refer to the [@sec-appendix].

## Variables Explanation
To better understand the data, the summary_table dataset was developed to offer a more structured and aggregated view of the crime data. By transforming and summarizing the data into a format that displays crime types against years, this dataset simplifies the task of identifying and analyzing trends over time. The dataset includes various crime categories, such as assault, theft, and other types, with each variable representing the annual count of reported cases for each crime type. This longitudinal data provides a detailed view of the crime trends in Toronto over a decade. 
```{r}
#| label: tbl-data-summary
#| tbl-cap: Crime Incidents Summary by Year
#| echo: false

# Read the data
summary_table <- read.csv(here("outputs/data/summary_table.csv"), check.names = FALSE)

@@

# Read the data
two_types_crime <- read.csv(here("outputs/data/two_types_crime.csv"), check.names = FALSE)

@@

# Plotting the line graph
ggplot(long_data, aes(x = Year, y = Count, group = Crime_Type, color = Crime_Type)) +
  geom_line(linewidth = 1) + # Use linewidth instead of size
  geom_point(size = 1.5) + # Point size
  theme_minimal() + # Minimal theme
  labs(title = "Crime Count by Type Over Years",
         x = "Year",
         y = "Number of Cases",
         color = "Crime Type") +
  theme(
    plot.title = element_text(hjust = 0.5), # Center the plot title
    legend.title = element_blank(), # Remove the legend title
    axis.text.x = element_text(angle = 45, hjust = 1), # Angle the x-axis text for readability
    text = element_text(size = 16), # Increase the global font size
  )
```
@fig-trends-in-crime reveals distinct trends for various crime types in Toronto over the years. We notice that assault crimes have been on the rise in recent years. This could signal a growing concern in public safety and might prompt law enforcement to investigate the underlying causes, such as increased urbanization, changes in population density, or other socio-economic factors. Conversely, there is a noticeable decline in theft-related crimes, which could reflect the success of recent crime prevention initiatives, improved security measures, or increased public awareness.

@@

* **Violent Crimes (Red Line):** There has been an overall increasing trend in the number of violent crime cases over the years. There are some fluctuations, with a noticeable dip occurring in one of the years, but the general direction is upwards, especially towards the end of the period, indicating a rise in violent crimes or an increase in reporting.

@@

## Weaknesses and Next Steps

While the visualization provides a high-level overview of crime trends in Toronto, it's essential to conduct a more detailed analysis to understand the underlying causes of these trends. This includes examining subcategories of crimes, focusing on specific geographic areas within the city, and considering other relevant data sources.

Furthermore, this study emphasises the need for ongoing monitoring and analysis of crime data to comprehend the ever-changing dynamics of urban crime. In light of these findings, future studies should focus on understanding the influence of specific policies and social changes on crime rates. Only by doing such a comprehensive analysis can we hope to develop strategies that will ensure the safety and well-being of all Toronto residents.
